Quality Definitions of Qualities

### 2021-09-22 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
Edits to quality paper - mostly changes and additions to the introduction

\documentclass[letterpaper,cleveref]{lipics-v2019}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\hypersetup{
colorlinks=true,        % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with linkbordercolor)
citecolor=blue,         % color of links to bibliography
filecolor=magenta,      % color of file links
urlcolor=cyan           % color of external links
}

@@

Given that our research focuses on improving the quality of software, we
often find it necessary to have clear definitions of the different software
qualities, like maintainability, usability, sustainability, etc.
Unfortunately, the definitions that we find in the literature are not always
satisfying.  This became clear when we recently undertook an exercise for
measuring the \lq\lq{}state of the practice\rq\rq{} in different software
domains, including the domains of seismology software \citep{SmithEtAl2018},
Geographic Information Systems \citep{SmithEtAl2018_arXivGIS} and mesh
generators \citep{smith2016state}. In trying to determine methods to quickly
assess quality of large samples of existing software, we found communication
between ourselves and other challenging.  We craved unambigous and
standardized definitions to enable methods for measuring qualities.
Therefore, we have endeavored to collect an exhaustive list of definitions
and from this build a set of recommended definitions, based on our quality
criteria for high-quality quality definitions.  We are publishing our
collection, since others may likewise find them valuable.

@@

Not specifically emphasizing scientific computing, but when a specific
context is needed for a definition, the scientific context will be used.  For
instance, the ISO definition of accuracy is defined in terms of precision, but
in scientific computing accuracy and precision are two different concepts.

@@

\subsection{Background}

@@

\begin{defn}
\textsf{Context of use}:
characteristics of the users, tasks and the organizational and physical
environments 
\cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Goal}: intended outcome
\cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Task}: activities required to achieve a goal 
\cite{ISO9241:11}
\end{defn}

>>>>

\documentclass[letterpaper, cleveref]{lipics-v2019}

\usepackage{natbib}
\usepackage{booktabs}
\usepackage{amsmath,amsthm}
\usepackage{hyperref}
\hypersetup{
colorlinks=true,        % false: boxed links; true: colored links
linkcolor=red,          % color of internal links (change box color with linkbordercolor)
citecolor=blue,         % color of links to bibliography
filecolor=magenta,      % color of file links
urlcolor=cyan           % color of external links
}
\usepackage{bbding}

@@

In the course of conducting and presenting our software engineering related
research we have often desired clear definitions for relevant software
qualities, like maintainability, usability, productivity, sustainability, etc.
Unfortunately, the definitions of qualities that we found in the literature were
not always satisfying.  The inadequacy of existing definitions became
particularly clear when we recently undertook an exercise measuring the
\lq\lq{}state of the practice\rq\rq{} in different software domains, including
the domains of seismology software \citep{SmithEtAl2018}, Geographic Information
Systems \citep{SmithEtAl2018_arXivGIS} and mesh generators
\citep{smith2016state}. When determining methods to quickly assess quality
across large samples of existing software, we found communication amongst
ourselves and with others challenging. We craved unambiguous and standardized
definitions to enable and inspire methods for measuring software qualities.  To
help ground our future research, and possibly help others in the same situation,
we have collected an exhaustive list of definitions.  From this list we have
produced our set of recommended software quality definitions.

In the current work, quality is defined as a measure of the excellence or worth
of an entity.  As is standard practice, we do not think of quality as a single
measure, but rather as a set of measures.  In the overview of managing software
quality provided by \citet[p 110]{VanVliet2000} the different measures are
called quality attributes, quality factors, and quality criteria.  In
\citet{ISO9126} they are called quality characteristics, subcharacteristics and
attributes.  To keep our presentation simple, and to not imply a classification
hierarchy that we don't intend, we will just use the term qualities to refer to
the different measures of quality.

Many taxonomies have been proposed for software qualities, as in
\citet{McCallEtAl1977, boehm1976quantitative, ISO9126}. Rather than aiming for a
complete coverage of the history of qualities, the wide variety of potential
qualities, or the relationships between them, we will instead take a practical
approach; we will focus on the qualities that we have found most relevant in our
research.  This means we will include the usual software engineering suspects,
like usability and maintainability.  Furthermore, since we also do research on
scientific computing software, we will also include scientifically relevant
qualities, like validatability and reproducibility relevant.

Given our respect for \citet[p.\ 15--33]{GhezziEtAl2003}, we used their list of
qualities as a starting point. Our list, shown in
Table~\ref{Tab_SoftwareQualities}, includes all of their qualities, except for
timeliness, which we excluded because the already included quality of
productivity implies the importance of timeliness.  To the list from
\citet{GhezziEtAl2003} we have added the practical quality of installability,
since our aforementioned studies on the state of the practice in different
software domains (for instance \citet{SmithEtAl2018_arXivGIS}) have shown us
that none of the other qualities matter if the software cannot be properly
installed.  Given our interest in scientific computing, we also added the
qualities of validatability and reproducibility. Since we were not able to find
much information in the literature on the quality of verifiability, we also
added the closely related quality of testability to our List.  The quality of
sustainability was added since this term has emerged as a popular goal for
software projects. The order of the qualities in
Table~\ref{Tab_SoftwareQualities} follow the order used in
\cite{GhezziEtAl2003}, which puts related qualities (like correctness and
reliability) together.  The order is roughly the order in which qualities are
considered.  We followed these same guidelines when inserting our additional
qualities into the list.

\begin{table}[h!]
  \begin{center}
  \begin{tabular}{lllll}
  \toprule
  & External & Internal & Product & Process \\  
  \midrule
  Installability & \Checkmark & & \Checkmark & \\
  Correctness & \Checkmark & \Checkmark & \Checkmark & \Checkmark \\
  Reliability & \Checkmark & & \Checkmark & \Checkmark \\
  Robustness & \Checkmark & & \Checkmark & \Checkmark \\     
  Performance & \Checkmark & & \Checkmark & \\
  Usability & \Checkmark & & \Checkmark & \Checkmark \\
  Verifiability & & \Checkmark & \Checkmark & \Checkmark \\
  Validatability & \Checkmark & & \Checkmark & \\     
  Testability & & \Checkmark & \Checkmark & \\
  Maintainability &  & \Checkmark & \Checkmark & \Checkmark\\
  Reusability & & \Checkmark & \Checkmark & \Checkmark\\
  Portability & & \Checkmark & \Checkmark & \\
  Understandability & & \Checkmark &  & \\     
  Interoperability & \Checkmark & & \Checkmark & \\
  Productivity &  &  &  & \Checkmark \\
  Visibility &  & &  & \Checkmark \\
  Reproducibility & \Checkmark & & \Checkmark & \\
  Sustainability & & \Checkmark & \Checkmark & \\      
  \bottomrule
\end{tabular}
\end{center}
\caption{Software Qualities}
\label{Tab_SoftwareQualities}
\end{table}

Table~\ref{Tab_SoftwareQualities} makes a distinction between external and
internal qualities.  External qualities are the qualities that are relevant to
the user of the software while internal qualities are those that are relevant to
developers \citep[p.\ 16]{GhezziEtAl2003}.  Although the users are not typically
interested in internal qualities, it is the internal qualities that influence
external quality.  External qualities are assessed using the produced software
and any user relevant documentation.  Internal qualities reference such things
as the code, documentation, test cases, and the software development process.
Correctness is shown as both an internal and external quality because it is
possible to judge the correctness of software and of the various internal
artifacts produced while developing software.  The external quality of
reliability is influenced by the internal quality of verifiability.  Since
validatability is related to getting the right requirements, it is shown as a
external quality.  Although one could judge the usability of internal
documentation, the quality of usability is considered as external, with the
quality of understandability introduced to cover the internal equivalent.
Filling in the table the way we want to use it.

Some qualities focus on the product, others on the process, and yet others on
both.  This is shown in Table~\ref{Tab_SoftwareQualities}.  Define process.
Define product. Process quality improves product quality.  The check marks are
for qualities that can be applied to a process, not the qualities that a process
improves.  Examples.  Performance isn't for process.  For process use
productivity.

@@

The topic of quality has been visited and revisited many times through the
course of software engineering.  Starting with McCall, Boehm.  Iso standards.
Taxonomies, relationships between qualities.  Rather than introduce distinctions
that aren't relevant, we'll just focus on qualities of interest. We are not
endeavoring to review all of the available literature, nor are we trying to
provide a complete framework for all qualities and the relationships between
them.  We are taking a more practical approach.  We have identified the
qualities that come up most in our research.  Not specifically emphasizing
scientific computing, but when a specific context is needed for a definition,
the scientific context will be used.  For instance, the ISO definition of
accuracy is defined in terms of precision, but in scientific computing accuracy
and precision are two different concepts.  Some are product and some are process
and some are both.  Some are internal and some are external.

@@

Roadmap.
\subsection{Background}

@@

\begin{defn}
\textsf{Context of use}:
characteristics of the users, tasks and the organizational and physical
environments. \cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Goal}: intended outcome. \cite{ISO9241:11}
\end{defn}

\begin{defn}
\textsf{Task}: activities required to achieve a goal. \cite{ISO9241:11}
\end{defn}
